---
title: "p8180_hw2_jj3205"
author: "Jia Ji (jj3205)"
date: "2022-10-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1


```{r}

library(tidyverse)
library(readxl)
```

Below we import and clean data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. The process begins with data import, updates variable names, and selects the columns that will be used in later parts fo this problem. We update `entry` from `yes` / `no` to a logical variable. As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

As it stands, these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 

The following code chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

The next code chunk is similar, but filters according to ADA compliance as an initial step. This produces a dataframe in which the number of rows is the number of ADA compliant stations. 

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```


## Problem 2


```{r, echo=FALSE}

library("readxl")

mrwheel = read_excel("data/Trash_Wheel_Collection_Data.xlsx", 
           sheet = "Mr. Trash Wheel", skip = 1,
           range = cellranger::cell_cols(("A:N")), ) %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  filter(dumpster != "Grand Total") %>%
  mutate(sports_balls = as.integer(round(sports_balls,0)),
         dumpster = as.integer(dumpster)) %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards)
view(mrwheel)
```


```{r}

profwheel = read_excel("data/Trash_Wheel_Collection_Data.xlsx",
           sheet = "Professor Trash Wheel", skip = 1,
           range = cellranger::cell_cols(("A:N"))) %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  dplyr::select(-x14)

view(profwheel)

mr_tidy =  mutate(mrwheel, Trash_wheel = c('Mr'), dumpster = as.character(dumpster))
prof_tidy = mutate(profwheel,
         Trash_wheel = c('Pro'),
         dumpster = as.character(dumpster),
         year = as.character(year))
wheel_tidy =  bind_rows(mr_tidy, prof_tidy) %>% 
  relocate(Trash_wheel, .before = "weight")


view(wheel_tidy)

total_weight = sum(profwheel$weight) 
total_ball = mrwheel %>% 
  filter(year == "2020") %>% 
  pull(sports_balls) %>% 
  sum() 

```

Total weight of trash collected by Professor Trash Wheel is `r total_ball`


## Problem 3


###Clean the data in pols-month.csv

```{r}
Pols_month_df = read_csv(
  "./data/pols-month.csv") %>%
  janitor::clean_names() %>% 
  drop_na() %>% 
  separate(mon, c("year","month","day")) %>%
  mutate(
    year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%
  mutate(month = month.abb[month]) %>% 
  mutate(president = ifelse(prez_gop == "0", "dem","gop")) %>% 
  relocate(president, .before = "gov_gop") %>% 
  dplyr::select(-prez_dem, -prez_gop, -day)
```

```{r}
view(Pols_month_df)
```

###Clean the data in snp.csv

```{r}
snp_df = read_csv(
  "./data/snp.csv") %>%
  janitor::clean_names() %>%
  janitor::clean_names() %>% 
  mutate(
    date = lubridate::parse_date_time2(date,orders = "mdy", cutoff_2000 = 23)) %>% 
  separate(date, c("year","month","day")) %>%
  mutate(
    year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%
  mutate(month = month.abb[month]) %>% 
  relocate(year, .before = "month") %>% 
  dplyr::select(-day) 
```

```{r}
view(snp_df)
```

###Clean the data in unemployment.csv

```{r}
unemploy_df = read_csv(
  "./data/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,names_to = "month", values_to = "unemployment_rate") %>%
  janitor::clean_names() 
```

```{r}
view(unemploy_df)
```

###Join the datasets

```{r}
snp_pols = 
  left_join(Pols_month_df, snp_df, by = c("year", "month")) 

merge_df = 
  left_join(snp_pols, unemploy_df, by = c("year", "month")) %>% 
View()


```
